---
title: "Tidyverse BMI 704 Laboratory"
author: "Chirag Patel and Chirag Lakhani"
date: "1/21/2018"
output: html_document
---

```{r setup, include=FALSE}
#install.packages('DT')
library(DT)
library(tidyverse)
library(broom)
knitr::opts_chunk$set(echo = TRUE)
load('nh_bmi704.Rdata')
```

# Data Description
These data are from the National Health and Nutrition Examination Survey (NHANES), a non-institutionalized sampling of the United States: [https://www.cdc.gov/nchs/nhanes/index.htm](NHANES)

In this lab we will explore phenotypes (P) of the metabolic syndrome, a risk factor for type 2 diabetes and heart disease.

# Data catalog

```{r}
DT::datatable(VarDescription)
```



# Data Manipulation

* create a new variable called *gender* (using the *RIAGENDR* column) with values and append to the dataframe **nhData**
    + M - males
    + F - females
  
* create a new variable called *ethnicity* (using the *RIDRETH1* column) with the following values and append to the dataframe **nhData**
    + H - Mexican American or Other Hispanic
    + W - White
    + B - Black
    + O - Other

* create a new variable called *age_group* (using the *RIDAGEYR* column) with the following values and append to the dataframe **nhData**
    + [0-18)
    + [18-30)
    + [30-60)
    + [60+)
    

* create a new variable called *obesity* (using the *BMXBMI* column) with the following values and append to the dataframe **nhData**
    + normal - if BMXBMI < 30
    + obese - if BMXBMI >= 30



* create a new variable called *diabetes* (using the *LBXGLU* column) with the following values and append to the dataframe **nhData**
    + normal - if LBXGLU < 126.0
    + diabetic - if LBXGLU >= 126.0
    

```{r}

nhData <- nhData %>% mutate(gender = case_when(RIAGENDR == '1' ~ 'M', 
                                               RIAGENDR == '2' ~ 'F' ))


nhData <- nhData %>% mutate(ethnicity = case_when(RIDRETH1 == '1' | RIDRETH1 == '2'  ~ 'H', 
                                                  RIDRETH1 == '3' ~ 'W',
                                                  RIDRETH1 == '4' ~ 'B',
                                                  RIDRETH1 == '5' ~ 'O',))



nhData <- nhData %>% mutate(age_group = case_when(RIDAGEYR >= 0 & RIDAGEYR < 18   ~ '[0,18)', 
                                                  RIDAGEYR >= 18 & RIDAGEYR < 30   ~ '[18,30)',
                                                  RIDAGEYR >= 30 & RIDAGEYR < 60   ~ '[30,60)',
                                                  RIDAGEYR >= 60   ~ '[60+)',))



nhData <- nhData %>% mutate(obesity = case_when(BMXBMI < 30   ~ 'normal', 
                                                BMXBMI >= 30   ~ 'obese'))
                                                

nhData <- nhData %>% mutate(diabetes = case_when(LBXGLU < 126.0   ~ 'normal', 
                                                LBXGLU >= 126.0   ~ 'diabetic'))
                                                


```










# Data Exploration with dplyr


### Find the following counts

* Number of individuals by gender
* Number of individuals by ethnicity
* Number of individuals by age group
* Number of individuals by gender and ethnicity (order by ethnicity and then gender)

```{r}

nhData %>% group_by(gender) %>% summarise(numIndividuals=n())

nhData %>% group_by(ethnicity) %>% summarise(numIndividuals=n())

nhData %>% group_by(age_group) %>% summarise(numIndividuals=n())


nhData %>% group_by(gender,ethnicity) %>% summarise(numIndividuals=n()) %>% arrange(ethnicity,gender)


```



### More practice with dplyr

* Create a dataframe called **bmi_ldl_df_age_gender** which contains only BMI, LDL, age group, and gender
* Using **bmi_ldl_df_age_gender** create a new dataset called **bmi_ldl_age_female** which only contains females
* Using the data frame **bmi_ldl_age_female** estimate the mean bmi and ldl level for each age group


```{r}
bmi_ldl_df_age_gender <- nhData %>% select(BMXBMI,LBDLDL,age_group,gender)

bmi_ldl_df_age_female <- bmi_ldl_df_age_gender %>% filter(gender =='F')

bmi_ldl_df_age_female %>% group_by(age_group) %>% summarise(avgLDL=mean(LBDLDL), avgBMI=mean(BMXBMI))


```
  
  
  
# Plotting with ggplot2


* What is the distribution of age in the population? Use 60 bins instead of default of 30.
```{r}

ggplot(nhData,aes(RIDAGEYR)) + geom_histogram(bins=60)


```


 Create boxplots of BMI, systolic blood pressure and LDL by gender (using the **nhData** dataframe) (give each gender a different color)


```{r}

ggplot(nhData, aes(x=gender, y=BMXBMI, fill=gender)) + geom_boxplot() 
ggplot(nhData, aes(x=gender, y=MSYS, fill=gender)) + geom_boxplot() 
ggplot(nhData, aes(x=gender, y=LBDLDL, fill=gender)) + geom_boxplot() 


```


* Create boxplots of BMI, HDL and LDL by gender that are facetted by ethnicity.

```{r}
ggplot(nhData, aes(x=gender, y=BMXBMI, fill=gender)) + geom_boxplot() + facet_wrap(~ethnicity) + scale_y_log10()
ggplot(nhData, aes(x=gender, y=MSYS, fill=gender)) + geom_boxplot() + facet_wrap(~ethnicity) + scale_y_log10()
ggplot(nhData, aes(x=gender, y=LBDLDL, fill=gender)) + geom_boxplot() + facet_wrap(~ethnicity) + scale_y_log10()



```


* Create a boxplot of systolic blood pressure by type 2 diabetes status facetted by gender (color by diabetes status).

[http://diabetes.diabetesjournals.org/content/66/2/543](See Mendelian Randomization Study relating systolic blood pressure to T2D)

```{r}


ggplot(nhData, aes(x=diabetes, y=MSYS, fill=diabetes)) + geom_boxplot() + facet_wrap(~gender) 


```


















# Correlations
- Repeat above correlations with remaining P: glucose,HDL, LDL
- Plot the correlations in different age, sex, and race/ethnciicity groups

* Find the correlation between age and BMI


```{r}

nhData %>% summarise(cor(RIDAGEYR,BMXBMI))




```

* Calculate the correlation between age and BMI and count the number of people for each gender
```{r}

nhData %>% group_by(gender) %>% summarise(cor(RIDAGEYR,BMXBMI), numberpeople=n())

```


* 






# Linear Modeling
- correlate each variable with each other
- create a linear model predicting each P as a function of age, sex, race/ethncity
- what is the R2 of each?
- what is the coeffcicients of the lienar model for each risk factor (age, sex, race/ethncitiy)? What is the interpetation?



* create a linear model predicting BMI as a function of age, sex, race/ethncity, LDL cholestrol, Triglycerides, and Systolic Blood Pressure and output the coefficients into a data frame called **fit_tidy** and the output model summary into a data frame called **fit_glance**, what is the $R^2$


```{r}

regressions <-  nhData %>% do(fit=lm(BMXBMI ~ RIDAGEYR+ gender+ ethnicity + LBDLDL+MSYS + LBXTR, data = .))



fit_tidy <- regressions %>% tidy(fit)

fit_glance <- regressions %>% glance(fit)

DT::datatable(fit_tidy)


DT::datatable(fit_glance)



```



